Xor-Some Preregistered Study: Analyses

### 2021-10-18 analysis/prereg/03_xor-some_prereg_analysis.Rmd
refactor prereg R scripts

---
title: "Xor & Some Final analysis"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

@@

## Read Data

First, the preprocessed and tidy wide pilot data is read and combined. It comprises data from N = 140 subjects after exclusions in total.
```{r, echo=FALSE}
pilot1 <- read_csv("./../data/pilots/pilot2_critical_zScore_wide_tidy.csv") %>%
  select(-competence_wUtt, -relevance_wUtt)
pilot2 <- read_csv("./../data/pilots/pilot3_critical_zScore_wide_tidy.csv")

d_both <- pilot1 %>% mutate(pilot = 1) %>% 
  rbind(., pilot2 %>% mutate(pilot = 2)
        )
# the data read here should be tidy, in correct format, bad workers should be excluded already
# z-scored data should be used and simulated!

# separate some and xor data!
d_both_some <- d_both %>% filter(main_type == "some") 
d_both_xor  <- d_both %>% filter(main_type == "xor")
```
Sanity check the combined data (as can be seen, z-scored data is used):
```{r, echo=FALSE, message=FALSE, warning=FALSE}
d_both %>%
  ggplot(., aes(x = relevance, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("Inference strength ratings") +
  facet_wrap(~main_type, ncol = 1) -> p.rel

d_both %>%
  ggplot(., aes(x = competence, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.comp

d_both %>%
  ggplot(., aes(x = prior, y = target)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_smooth(method = "lm") +
  ylab("") +
  facet_wrap(~main_type, ncol = 1) -> p.pri

gridExtra::grid.arrange(p.rel, p.comp, p.pri, ncol = 3) 
```

## Model

Next, fit the maximal model *across* experiments (i.e., *across SI trigger types*). For convergence reasons, correlation of random effects has to be set to 0. 
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
d_both <- d_both %>% mutate(
  main_type = as.factor(main_type)
)
# some = 0, xor = 1
contrasts(d_both$main_type) 
# try to fit maximal model with brm
model_SI <- brm(target ~ prior*competence*relevance* main_type +
                   (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                 data = d_both,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 3000)
```

@@

## Extract conrtrasts of interest

@@

## Perform BF analysis

@@

```{r}
# tiny update for separating the trigger words 
test_conjunction_of_all_hypotheses2 <-  function(posterior_hypotheses, trigger) {
  posterior_hypotheses %>% 
    mutate(hypothesis_true = case_when(
      key == paste('competence_', trigger, sep='') ~ positive_eff > 0.95,
      key == paste('prior_', trigger, sep='') ~ negative_eff > 0.95,
      key == paste('relevance_', trigger, sep='') ~ positive_eff > 0.95
    )) %>% 
    pull(hypothesis_true) %>% all()
}
# applied to the pilot data
test_conjunction_of_all_hypotheses2(posterior_hypotheses, "xor")
test_conjunction_of_all_hypotheses2(posterior_hypotheses, "some")
```

## Experimenting with multiple regression

```{r}
dummy_data <- tribble(
  ~main, ~REL, ~target,
  "some", 0, 4,
  "some", 1, 6,
  "xor", 0, 0,
  "xor", 1, 12
)
# check that the slope of "REL" for "xor" is 12
# this is calculated as betat_REL + beta_mainxor:REL
# importantly: the intercept adjustment beta_mainxor doesn't come in!
lm(target ~ main * REL, dummy_data)
```

>>>>

---
title: "Xor-Some Preregistered Study: Analyses"
author: "Polina Tsvilodub"
date: "8/31/2021"
output: github_document
---

@@

Read data created in preprocessing script:
```{r}
d_critical_zScored_wide <- read_csv("./../../data/results_prereg_tidy_zScored_wide.csv")
d_critical_zScored <- read_csv("./../../data/results_prereg_tidy_zScored_long.csv")
```

## Main analysis

Next, fit the maximal model *across* experiments (i.e., *across SI trigger types*). For convergence reasons, correlation of random effects has to be set to 0. 
```{r, results='hide', message=FALSE, warning=FALSE}
# Full model 
d_critical_zScored_wide <- d_critical_zScored_wide %>% mutate(
  main_type = as.factor(main_type)
)
# some = 0, xor = 1
contrasts(d_critical_zScored_wide$main_type)

# set priors
priors <- set_prior("student_t(1, 0, 2)", class = "b")

model_SI <- brm(target ~ prior*competence*relevance* main_type +
                  # (1 + prior + competence + relevance + main_type || submission_id) +
                   (1 | title),
                 data = d_critical_zScored_wide,
                 prior = priors,
                 sample_prior = T,
                 control = list(adapt_delta = 0.95),
                 cores = 4,
                 iter = 3000)
```

@@

@@

#### Perform BF analysis

@@

## Exploratory analyses

Model with binary predictors:
```{r}
d_critical_zScored %>% 
  select(submission_id, title, main_type, block_extended, response_centered, prior_class, class_condition) %>% 
  unique() %>% 
  pivot_wider(
    names_from = class_condition, 
    values_from = prior_class
  )  %>%
  mutate(relevance = factor(relevance, levels = c(1, 0)),
         competence = factor(competence, levels = c(1, 0)),
         prior = factor(prior, levels = c(1, 0)),
         main_type = as.factor(main_type)) %>% 
  filter(block_extended == "target") -> d_critical_wide_cat_zScored
# sum code predictors
# low prior: -1, high prior:: 1 
contrasts(d_critical_wide_cat_zScored$prior) <- contr.sum(2)
# low comp : -1, high comp : 1
contrasts(d_critical_wide_cat_zScored$competence) <- contr.sum(2)
# low rel: -1, high rel: 1
contrasts(d_critical_wide_cat_zScored$relevance) <- contr.sum(2)
# leave main type dummy coded
contrasts(d_critical_wide_cat_zScored$main_type)

# xor, maximal model with interactions and maximal REs
model_xor_cat_zScored <- brm(
  response_centered ~ prior*competence*relevance*main_type + 
    (1 + prior + competence + relevance + main_type || submission_id) +
    (1 | title),
  data = d_critical_wide_cat_zScored,
  prior = priors,
  sample_prior = T,
  family = "gaussian",
  cores = 4,
  iter = 3000
)

```
```{r}
summary(model_xor_cat_zScored)
```

Model with aggregated by-story ratings:
```{r, results='hide'}
d_critical_zScored_wide_byItem <- d_critical_zScored_wide %>% ungroup() %>%
  group_by(title, main_type) %>%
  summarize(mean_response = mean(target),
            mean_prior = mean(prior),
            mean_relevance = mean(relevance),
            mean_competence = mean(competence))

model_means <- brm(mean_response ~ mean_prior + mean_relevance + mean_competence + main_type,
                       data = d_critical_zScored_wide_byItem,
                       prior = priors,
                       sample_prior = T,
                       family = "gaussian",
                       cores = 4,
                       iter = 3000
                       )
```

```{r}
summary(model_means)
```
